---
title: "E2S_KDdataAnalysis"
author: "Muriel Heldring"
date: '2022-08-31'
output: html_document
---

First load packages:
```{r}
library(tidyverse)
library(naniar)
library(splines)
library(dplyr)
```

```{r, setup, include=FALSE}
rm(list = ls())

knitr::opts_knit$set(root.dir = "/data/muriel/Projects/E2S/Analysis/00_Imaging/Exp013_GFP_Britt/")
```

Load data sets: (5 KD EXPs)
```{r}
dir("Data_KD")
KD_006MT <- read_delim("Data_KD/006-MT/20200224_Exp006-MT_t24_SummaryData.txt", delim = "\t")
KD_252 <- read_delim("Data_KD/252/20200612_Exp252_Nikon4_live_Summarydata.txt", delim = "\t")
KD_262 <- read_delim("Data_KD/262/20200612_Exp262_Nikon2_siRNAKD_Summarydata.txt", delim = "\t")
KD_267 <- read_delim("Data_KD/267/20200627_Exp267_24hrsliveKD_Summarydata.txt", delim = "\t")
KD_273 <- read_delim("Data_KD/273/20200704_Exp273_Nikon4_24hrsliveKD_SummaryData.txt", delim = "\t")
```

Convert tables to wide format:
```{r}
KD_006MT_wider <- pivot_wider(data = KD_006MT, 
                    id_cols = c("treatment","plateID","dose_uM","timeID","locationID", "cell_line"),
                    names_from = "variable",
                    values_from = "value")
KD_252_wider <- pivot_wider(data = KD_252, 
                    id_cols = c("treatment","plateID","dose_uM","timeID","locationID", "cell_line"),
                    names_from = "variable",
                    values_from = "value")
KD_262_wider <- pivot_wider(data = KD_262, 
                    id_cols = c("treatment","plateID","dose_uM","timeID","locationID", "cell_line"),
                    names_from = "variable",
                    values_from = "value")
KD_267_wider <- pivot_wider(data = KD_267, 
                    id_cols = c("treatment","plateID","dose_uM","timeID","locationID", "cell_line"),
                    names_from = "variable",
                    values_from = "value")
KD_273_wider <- pivot_wider(data = KD_273, 
                    id_cols = c("treatment","plateID","dose_uM","timeID","locationID", "cell_line"),
                    names_from = "variable",
                    values_from = "value")
```

Needed KD_data:
	Knock down data of proteins and its effect on expression in other proteins
	Five experiments, and use for
	- GREB1-GFP: Exp. 006-MT, Exp. 252 , Exp. 262 en Exp. 273
	- PGR-GFP: Exp. 006-MT, Exp. 262 en Exp. 267
	- TFF1-GFP: Exp. 006-MT, Exp. 262 en Exp. 273
	
Remove unused data
```{r}
KD_252_wider_GREB1 <- KD_252_wider %>% filter(cell_line %in% c("GREB1-GFP_mock","GREB1-GFP_siESR1", "GREB1-GFP_siGREB1", "GREB1-GFP_siKP", "GREB1-GFP_siPGR", "GREB1-GFP_siTFF1", "GREB1-GFP_untreated", "GREB1-GFP_medium"))

##same variable names
KD_252_wider_GREB1 <- KD_252_wider_GREB1 %>% rename(Cytoplasm_Intensity_IntegratedIntensity_Image_GFP = Cytoplasm_N_Intensity_IntegratedIntensity_Image_GFP)

KD_267_wider_PGR <- KD_267_wider %>% filter(cell_line %in% c("PGR-GFP_mock", "PGR-GFP_siESR1", "PGR-GFP_siGREB1", "PGR-GFP_siPGR", "PGR-GFP_siTFF1", "PGR-GFP_siKP", "PGR-GFP_untreated", "PGR-GFP_medium"))

KD_273_wider_GREB1_TFF1 <- KD_273_wider %>% filter(cell_line %in% c("GREB1-GFP_mock", "GREB1-GFP_siESR1", "GREB1-GFP_siGREB1", "GREB1-GFP_siPGR", "GREB1-GFP_siTFF1", "GREB1-GFP_siKP", "GREB1-GFP_untreated","GREB1-GFP_medium", "TFF1-GFP_mock", "TFF1-GFP_siESR1", "TFF1-GFP_siGREB1", "TFF1-GFP_siPGR", "TFF1-GFP_siTFF1", "TFF1-GFP_siKP", "TFF1-GFP_untreated", "TFF1-GFP_medium"))

##rename columns
KD_006MT_wider <- KD_006MT_wider %>% rename(Cytoplasm_Intensity_IntegratedIntensity_Image_GFP = Cytoplasm_N_Intensity_IntegratedIntensity_Image_GFP)

KD_262_wider <- KD_262_wider %>% rename(Cytoplasm_Intensity_IntegratedIntensity_Image_GFP = Cytoplasm_N_Intensity_IntegratedIntensity_Image_GFP)

```

Knock down data:
```{r}
#bind data sets
KD_TOTAL <- bind_rows(KD_006MT_wider,
                      KD_252_wider_GREB1,
                      KD_262_wider,
                      KD_267_wider_PGR,
                      KD_273_wider_GREB1_TFF1)

rm(list = c("KD_006MT",
            "KD_006MT_wider",
            "KD_252",
            "KD_252_wider",
            "KD_252_wider_GREB1",
            "KD_262",
            "KD_262_wider",
            "KD_267",
            "KD_267_wider",
            "KD_267_wider_PGR",
            "KD_273",
            "KD_273_wider",
            "KD_273_wider_GREB1_TFF1"))

#rename DMSO to 17B-estradiol (DMSO = 0, E2 = 1)
KD_TOTAL <- KD_TOTAL %>% mutate(dose_uM = as.factor(dose_uM), 
                                 treatment = ifelse(treatment == "DMSO", "17B-estradiol",treatment))

#remove medium control
KD_TOTAL <- KD_TOTAL %>% filter(treatment !="Medium",
                                !cell_line %in% c("GREB1-GFP_untreated",
                                                  "TFF1-GFP_untreated",
                                                  "PGR-GFP_untreated",
                                                  "GREB1-GFP_medium",
                                                  "TFF1-GFP_medium",
                                                  "PGR-GFP_medium"))

# Rename the GFP intensity column names
KD_TOTAL <- KD_TOTAL %>% rename("NucleicIntensity" = "Nuclei_Intensity_IntegratedIntensity_Image_GFP",
                                "CytoplasmicIntensity" = "Cytoplasm_Intensity_IntegratedIntensity_Image_GFP")

```

```{r}
# Normalised cell count by dividing cell count with the cell count at timeID 0 
KD_TOTAL_TIME1 <- KD_TOTAL %>% filter(timeID == 1) %>% 
  select(-c("timeID", "NucleicIntensity",
            "CytoplasmicIntensity"))

KD_TOTAL <- left_join(KD_TOTAL, KD_TOTAL_TIME1, 
                      by = c("treatment", "plateID", "dose_uM", "locationID", "cell_line"), 
                      suffix = c("","_t1")) %>% 
                mutate(Relative_Cell_Growth = imageCountParentObj/imageCountParentObj_t1) %>%
                select(-c("imageCountParentObj_t1"))
rm("KD_TOTAL_TIME1")

# # Normalize intensities to DMSO
# KD_TOTAL_DMSO <- KD_TOTAL %>% filter(dose_uM == 0) %>%
#   select(-c(dose_uM, locationID, treatment)) %>% ungroup() %>%
#   group_by(cell_line, timeID, plateID) %>%
#   summarise(NucleicIntensity_DMSO = mean(NucleicIntensity),
#             CytoplasmicIntensity_DMSO = mean(CytoplasmicIntensity))
# 
# KD_TOTAL <- left_join(KD_TOTAL, KD_TOTAL_DMSO, by = c("cell_line", "timeID", "plateID"),
#                       suffix = c("", "_DMSO")) %>% 
#   mutate(RelativeNucleicIntensity = NucleicIntensity / NucleicIntensity_DMSO,
#          RelativeCytoplasmicIntensity = CytoplasmicIntensity / CytoplasmicIntensity_DMSO) %>% 
#   select(-c("NucleicIntensity_DMSO", "CytoplasmicIntensity_DMSO", "treatment"))
# rm("KD_TOTAL_DMSO")

```

Split the cell lines into 3:
```{r}
# Split cell line column 
KD_TOTAL <- KD_TOTAL %>%
  mutate(condition = sapply(cell_line, function(cl) {str_split(cl, "_")[[1]][2]}),
         cell_line = sapply(cell_line, function(cl) {str_split(cl, "-")[[1]][1]}),
         condition = ifelse(condition == "siKP", "siControl", condition))

```

Min-max normalisation
```{r}
#define function
min_max_norm <- function(x) {
    (x - min(x)) / (max(x) - min(x))
}


#create min max data set
KD_TOTAL_minmax <- KD_TOTAL %>% group_by(cell_line, plateID) %>% 
  mutate(NucleicIntensityMinMax = min_max_norm(NucleicIntensity),
         CytoplasmicIntensityMinMax = min_max_norm(CytoplasmicIntensity))
head(KD_TOTAL_minmax)

```


Technical Replicate Mean of KD data
```{r}

#calculate TechRepMean unnormalised and normalised:
TechRepMean_KD_minmax <- KD_TOTAL_minmax %>% ungroup() %>% 
  group_by(cell_line, timeID, plateID, condition, dose_uM) %>%
  summarize(#mean cyto intensity and sd
            MeanCytoplasmicIntensityMinMax = mean(CytoplasmicIntensityMinMax),
            sdCytoplasmicIntensityMinMax = sd(CytoplasmicIntensityMinMax, na.rm = T),
            #mean nucl intensity and sd
            MeanNucleicIntensityMinMax = mean(NucleicIntensityMinMax),
            sdNucleicIntensityMinMax = sd(NucleicIntensityMinMax, na.rm = T))
```

Biological Replicate Mean of KD data:
```{r}
#calculate BioRepMean unnormalised and normalised:
BioRepMean_KD_minmax <- TechRepMean_KD_minmax %>% ungroup() %>% 
  group_by(cell_line, timeID, condition, dose_uM) %>%
  # mean cyto intensity and sd
  summarise(BioMeanCytoplasmicIntensityMinMax = mean(MeanCytoplasmicIntensityMinMax),
            BioSDCytoplasmicIntensityMinMax = sd(MeanCytoplasmicIntensityMinMax, na.rm = T),
            # mean nucl intensity and sd
            BioMeanNucleicIntensityMinMax = mean(MeanNucleicIntensityMinMax),
            BioSDNucleicIntensityMinMax = sd(MeanNucleicIntensityMinMax, na.rm = T))
```

Biological relative mean intensity
```{r}
#Bio mean cytoplasmic intensity of GREB1
ggplot(BioRepMean_KD_minmax %>% 
         filter(dose_uM == 1, cell_line == "GREB1",
                !condition %in% c("siControl", "siESR1"))) +
  geom_errorbar(aes(x = timeID,
                    ymin = BioMeanCytoplasmicIntensityMinMax - BioSDCytoplasmicIntensityMinMax,
                    ymax = BioMeanCytoplasmicIntensityMinMax + BioSDCytoplasmicIntensityMinMax,
                    colour = condition), width = 0.5) +
  geom_line(aes(x= timeID, y = BioMeanCytoplasmicIntensityMinMax, color = condition), size = 1) + 
  labs(x = "Time (h)", y = "GFP intensity (a.u.)") + 
  scale_colour_viridis_d(name = "Condition") + 
  theme_classic() +
  #ylim(0,1.11) + 
  scale_y_continuous(breaks = c(0,0.2,0.4,0.6,0.8,1)) +
  theme(plot.title = element_text(hjust = 0.5, size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 18)) +
  ggtitle("GREB1-GFP")
ggsave(filename = "/data/muriel/Projects/E2S/Analysis/00_Imaging/Exp013_GFP_Britt/Figures/v2/Fig2C_KD_GREB1.pdf", width = 4.5, height = 3)

#Bio mean cytoplasmic intensity of TFF1
ggplot(BioRepMean_KD_minmax %>% 
         filter(dose_uM == 1, cell_line == "TFF1",
                !condition %in% c("siControl", "siESR1"))) +
  geom_errorbar(aes(x = timeID,
                    ymin = BioMeanCytoplasmicIntensityMinMax - BioSDCytoplasmicIntensityMinMax,
                    ymax = BioMeanCytoplasmicIntensityMinMax + BioSDCytoplasmicIntensityMinMax,
                    colour = condition), width = 0.5) +
  geom_line(aes(x= timeID, y = BioMeanCytoplasmicIntensityMinMax, color = condition), size = 1) + 
  labs(x = "Time (h)", y = "GFP intensity (a.u.)") + 
  scale_colour_viridis_d(name = "Condition") + 
  theme_classic() +
  ylim(0,1) +
  theme(plot.title = element_text(hjust = 0.5, size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 18)) +
  ggtitle("TFF1-GFP")
ggsave("/data/muriel/Projects/E2S/Analysis/00_Imaging/Exp013_GFP_Britt/Figures/v2/Fig2C_KD_TFF1.pdf", width = 4.5, height = 3)

#Bio mean nucleic intensity of PR
ggplot(BioRepMean_KD_minmax %>% 
         filter(dose_uM == 1, cell_line == "PGR",
                !condition %in% c("siControl", "siESR1"))) +
  geom_errorbar(aes(x = timeID,
                    ymin = BioMeanNucleicIntensityMinMax - BioSDNucleicIntensityMinMax,
                    ymax = BioMeanNucleicIntensityMinMax + BioSDNucleicIntensityMinMax,
                    colour = condition), width = 0.5) +
  geom_line(aes(x= timeID, y = BioMeanNucleicIntensityMinMax, color = condition), size = 1) + 
  labs(x = "Time (h)", y = "GFP intensity (a.u.)") + 
  scale_colour_viridis_d(name = "Condition") + 
  theme_classic() +
  ylim(0,1) +
  theme(plot.title = element_text(hjust = 0.5, size = 22),
        axis.text = element_text(size = 16),
        axis.title = element_text(size = 20),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 18)) +
  ggtitle("PR-GFP")
ggsave("/data/muriel/Projects/E2S/Analysis/00_Imaging/Exp013_GFP_Britt/Figures/v2/Fig2C_KD_PR.pdf", width = 4.5, height = 3)
```



Is the GFP expression between different conditions significant different from each other?
http://www.sthda.com/english/wiki/one-way-anova-test-in-r#compute-one-way-anova-test
```{r}
# Hypothesis: The mean GFP intensity of one condition within each cell line is not significantly different from the other condition within the same cell line


KD_data <- TechRepMean_KD_minmax %>% ungroup() %>% filter(timeID == 24, dose_uM == 1) %>% 
  select(cell_line, condition, plateID, MeanCytoplasmicIntensityMinMax, MeanNucleicIntensityMinMax) %>%
  mutate(Intensity = ifelse(cell_line %in% c("GREB1","TFF1"),
                            MeanCytoplasmicIntensityMinMax,
                            MeanNucleicIntensityMinMax))

# Make the variables categorical
KD_data <- KD_data %>% mutate(condition = as.factor(condition))

```

Do one way ANOVA per cell line:
```{r}
#PGR:
KD_data_PGR <- KD_data %>% filter(cell_line == "PGR")

group_by(KD_data_PGR, condition) %>%
  summarise(
    count = n(),
    mean = mean(Intensity, na.rm = TRUE),
    sd = sd(Intensity, na.rm = TRUE)
  )

# Compute the analysis of variance
res.aov_PGR <- aov(Intensity ~ condition, data = KD_data_PGR)
# Summary of the analysis
summary(res.aov_PGR)

#GREB1:
KD_data_GREB1 <- KD_data %>% filter(cell_line == "GREB1")
group_by(KD_data_GREB1, condition) %>%
  summarise(
    count = n(),
    mean = mean(Intensity, na.rm = TRUE),
    sd = sd(Intensity, na.rm = TRUE)
  )

# Compute the analysis of variance
res.aov_GREB1 <- aov(Intensity ~ condition, data = KD_data_GREB1)
# Summary of the analysis
summary(res.aov_GREB1)

#TFF1:
KD_data_TFF1 <- KD_data %>% filter(cell_line == "TFF1")
group_by(KD_data_TFF1, condition) %>%
  summarise(
    count = n(),
    mean = mean(Intensity, na.rm = TRUE),
    sd = sd(Intensity, na.rm = TRUE)
  )

# Compute the analysis of variance
res.aov_TFF1 <- aov(Intensity ~ condition, data = KD_data_TFF1)
# Summary of the analysis
summary(res.aov_TFF1)
```

Now we know that the conditions within the KD_data data set are significant different from each other in terms
of Real_GFP_intensity.
The next step is to check if there is sig dif between specific conditions.

Tukey multiple pairwise-comparisons:
As the ANOVA test is significant, we can compute Tukey HSD (Tukey Honest Significant Differences, R function:
TukeyHSD()) for performing multiple pairwise-comparison between the means of groups.
```{r}
TukeyHSD(res.aov_PGR)
#save("Significant difference between PGR condition.png")
TukeyHSD(res.aov_GREB1)
#save("Significant difference between GREB1 condition.png")
TukeyHSD(res.aov_TFF1)
#save("Significant difference between TFF1 condition.png")

```

